---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# MittagLeffleR

The MittagLeffleR `R` package calculates probabilities, quantiles and
random variables from both types Mittag-Leffler distributions. 

The first type Mittag-Leffler distribution is a heavy-tailed distribution, 
and occurs mainly as a waiting time distribution in problems with
"fractional" time scales, e.g. times between earthquakes.

The second type Mittag-Leffler distribution is light-tailed and, 
in a sense, inverse to the family of sum-stable distributions. 
It is used for time-changes of stochastic processes, to create 
"time-fractional" evolutions, e.g. anomalous diffusion processes.


## Installation

You can install MittagLeffleR from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("strakaps/MittagLeffler")
```

## Examples

### Fitting a Mittag-Leffler distribution to a random dataset

Fit a Mittag-Leffler distribution (first type) to a collection of
Mittag-Leffler random variables:

```{r, echo=TRUE}
library(MittagLeffleR)
y = rml(n = 100, tail = 0.9, scale = 2)
mlml <- function(X) {
  log_l <- function(theta) {
    #transform parameters so can do optimization unconstrained
    theta[1] <- 1/(1+exp(-theta[1]))
    theta[2] <- exp(theta[2])
    -sum(log(dml(X,theta[1],theta[2])))
  }
  ml_theta <- stats::optim(c(0.5,0.5), fn=log_l)$par
  #transform back
  ml_a <- 1/(1 + exp(-ml_theta[1]))
  ml_d <- exp(ml_theta[2])
  return(list("tail" = ml_a, "scale" = ml_d))
}
mlml(y)
```


### Calculate the probability density of an anomalous diffusion process

Brownian motion at time $t$ has a normal probability density $n(x,t^{1/2})$.
A fractional diffusion at time $t$ has the time-changed probability density 

$$p(x,t) = \int n(x,u)h(u,t)du$$

where $h(u,t)$ is a second type Mittag-Leffler probability density with 
scale $t^\alpha$:

```{r, warning=FALSE}
tail <- 0.65
dx <- 0.01
x <- seq(-3,3,dx)
umax <- qml(p = 0.99, tail = tail, scale = 1, second.type = TRUE)
u <- seq(0.01,umax,dx)
h <- dml(x = u, tail = tail, scale = 1, second.type = TRUE)
N <- outer(x,u,function(x,u){dnorm(x = x, sd = sqrt(u))})
p <- N %*% h * dx
plot(x,p, type='l', main = "Density of fractional diffusion at t=1")
```

